---
title: "Homework 7: Predator-Prey"
author: "Zach Loo"
execute:
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(sensitivity)
library(deSolve)
```

```{r}
source(here::here("7_predator_prey/lotvmodK.R"))
```

Set up initial model without hunting, given initial populations.

```{r}
# Initial populations
currpop <- c(prey = 40, pred = 1)

# Time range
days  <- seq(1, 250, by = 1)

# Parameters
parms <- c(rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000)

# Run ODE
results <- ode(func = lotvmodK, y = currpop, times = days, parms = parms)
```

```{r}
# Visualize
ggplot(data = results) +
    geom_line(aes(time, prey), color = 'darkgreen') +
    geom_line(aes(time, pred), color = 'red') +
    labs(x = "Days", y = "Count") +
    theme_light()
```

# Hunting Parms, Function Implement, & ODE

Goal for hunting implementation: Minimum threshold of prey must be met before hunting begins.

Prey-hunt threshold: 6?

```{r}
# Parameters
parms <- c(rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000, rhunt = 0.1, min_prey = 500)

# Model hunt implement
lotvmodKhunt <- function(t, pop, pars) {
  with(as.list(c(pars, pop)), {

    hunt  <- if (prey > min_prey) rhunt * prey else 0

    dprey <- rprey * (1 - prey / K) * prey - alpha * prey * pred - hunt
    dpred <- eff * alpha * prey * pred - pmort * pred
    return(list(c(dprey, dpred)))
  })
}
```

```{r}
# Initial populations
currpop <- c(prey = 1000, pred = 50)
# Time range
days  <- seq(1, 500, by = 1)
# Parameters
parms <- c(rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000, rhunt = 0.01, min_prey = 500)
results_hunt <- as.data.frame(ode(func = lotvmodKhunt, y = currpop, times = days, parms = parms))

# Visualize
ggplot(data = results_hunt) +
    geom_line(aes(time, prey), color = 'darkgreen') +
    geom_line(aes(time, pred), color = 'red') +
    labs(x = "Days", y = "Count") +
    theme_light()
```

# Uncertainty & Stability

```{r}
# --- Model ---
lotvmodKhunt <- function(t, pop, pars) {
  with(as.list(c(pars, pop)), {
    hunt  <- if (prey > min_prey) rhunt * prey else 0
    dprey <- rprey * (1 - prey / K) * prey - alpha * prey * pred - hunt
    dpred <- eff * alpha * prey * pred - pmort * pred
    return(list(c(dprey, dpred)))
  })
}

# --- Sampling ---
n_runs  <- 500
set.seed(42)
days    <- seq(1, 500, by = 1)
currpop <- c(prey = 1000, pred = 50)

param_samples <- tibble(
  rprey    = runif(n_runs, 0.95 * 0.8, 0.95 * 1.2),
  alpha    = runif(n_runs, 0.01 * 0.8, 0.01 * 1.2),
  eff      = runif(n_runs, 0.6  * 0.8, 0.6  * 1.2),
  pmort    = runif(n_runs, 0.4  * 0.8, 0.4  * 1.2),
  K        = 2000, 
  rhunt    = 0.1,       # change this & re-run
  min_prey = 500
)
```

```{r}
# --- Run simulations ---
run_model <- function(row) {
  p <- unlist(row)
  out <- tryCatch(
    as.data.frame(ode(func = lotvmodKhunt, y = currpop, times = days, parms = p)),
    error = function(e) NULL
  )
  if (is.null(out)) return(NA_real_)
  min(tail(out$prey, 250))
}

param_samples$min_prey_last250 <- map_dbl(1:n_runs, ~run_model(param_samples[.x, ]))
```

```{r}
# Stability flag based on min prey threshold
param_samples <- param_samples %>%
  mutate(stable = min_prey_last250 > 100)

# --- Summary ---
cat("Stable runs:", sum(param_samples$stable), "/", n_runs,
    "(", round(100 * mean(param_samples$stable), 1), "%)\n")
```

```{r}
# --- Plot: min prey distribution ---
ggplot(param_samples, aes(x = min_prey_last250, fill = stable)) +
  geom_histogram(bins = 40, color = "white", linewidth = 0.2) +
  geom_vline(xintercept = 100, linetype = "dashed", color = "orange", linewidth = 0.8) +
  scale_fill_manual(values = c("TRUE" = "#3fb950", "FALSE" = "#f85149"),
                    labels = c("TRUE" = "Stable", "FALSE" = "Unstable")) +
  labs(
    title    = paste0("Distribution of min prey (last 250d) across 500 random parameter sets  |  rhunt = ", unique(param_samples$rhunt)),
    subtitle = "Parameters varied ±20% around baseline. Orange dashed = stability threshold (min prey = 100).",
    x        = "Min prey population (days 250–500)",
    y        = "Count",
    fill     = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title     = element_text(face = "bold", size = 11),
    plot.subtitle  = element_text(color = "grey50", size = 9),
    legend.position = "bottom"
  )
```

I decided to define stability as if they prey population was over 100 during the last 250 days, this was mostly arbitrary and I genuinely tried to find a reasonable justification but playing around with the parameters was very difficult. Even when varying rhunt, I'd get results that did not make sense and would just give errors. I wish I could've done better with this assignment but I genuinely did not know how. For example, `rhunt` is currently 0.1 and if I increase this 0.5, I'd expect to get less stable runs, which they do but it goes from 2% to 0.2%, these values just seem oddly low.